---
title: "TEOS Hydrostation Data Lab"
author: "Juanita Carballeira"
date: "2023-03-07"
output: html_document
  #prettydoc::html_pretty:
    #theme: architect
   #highlight: github
---

## Load required libraries
```{r message = FALSE, warning = FALSE}
#install.packages("gws")
library(prettydoc)
library(tidyverse)
library(gsw)
library(readr)
```

## Now we need to import our data
```{r message = FALSE, warning = FALSE}
hydrostation_bottle <- read_delim("hydrostation_bottle.txt", 
    delim = "\t", escape_double = FALSE, 
    col_names = FALSE, trim_ws = TRUE, skip = 31)

hydrostation_bottle_names <- read_csv("hydrostation_bottle.txt", 
    skip = 30)

colnames(hydrostation_bottle) = colnames(hydrostation_bottle_names)
#view(hydrostation_bottle)
```

## Hydrostation S Discrete Bottle Data for years 1955 through December 2020.

### Variable Names and Units

- yyyymmdd = Year Month Day   
- decy   = Decimal Year     
- time   = Time (hhmm)      
- latN   = Latitude (Deg N) 
- lonW   = Longitude (Deg W)
- Depth  = Depth (m)                  
-Temp   = Temperature ITS-90 (C) 
- Pres   = CTD Pressure (dbar)   
- CTD_S  = CTD Salinity (PSS-78)      
- Sal1   = Salinity-1 (PSS-78)        
- Sig-th = Sigma-Theta (kg/m^3)       
- O2(1)  = Oxygen-1 (umol/kg)          
- OxFixT = Oxygen Fix Temp (C)        
- Anom1  = Oxy Anomaly-1 (umol/kg)    
Quality flags
- -999  = No data
- 0 = Less than detection limit
```{r}
# lets first plot the data
hydrostation_bottle %>% 
  filter(`Sig-th`!= -999) %>% # filter out -999 no data flag
  ggplot()+geom_point(aes(x=decy, y = `Sig-th`)) # hard to interpret even w no -999s

hydrostation_bottle %>% 
  filter(`Sig-th`!= -999 & Depth <20) %>% # filter out -999 no data flag and by upper 20m
  ggplot()+geom_line(aes(x=decy, y = `Sig-th`)) # line shows seasonality 
# clear seasonal signal for sigma-theta, lets see how this compares to temp
hydrostation_bottle %>% 
  filter(`Sig-th`!= -999 & Depth <20) %>% # filter out -999 no data flag and by upper 20m
  ggplot()+geom_point(aes(x=Temp, y = `Sig-th`))
```


```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```



```{r}
?gsw
?gsw_sigma0

?gsw_SA_from_SP
#practical salinity
#sea pressure dbar
#longitude
#latitude


##plot pressure data-missing before 1980s
hydrostation_bottle%>%
  ggplot()+geom_point(aes(x=decy,y=Pres))

#we have depth for time series
  
hydrostation_bottle %>%
  ggplot()+
  geom_point(aes(x=decy,y=Depth))

##Add pressure column
hydrostation_bottle=
  hydrostation_bottle %>%
  mutate(Pres_gsw=gsw_p_from_z(Depth*-1,latN))


  
hydrostation_bottle%>%
  ggplot()+
  geom_point(aes(x=decy, y=Sal1))

hydrostation_bottle=
  hydrostation_bottle%>%
  mutate(Pres_gsw=gsw_p_from_z(Depth*-1,latN))%>%
  mutate(s_abs_gsw=gsw_SA_from_SP(Sal1,Pres_gsw,360-lonW,latN))

#check it!
hydrostation_bottle%>%
  ggplot()+
  geom_point(aes(x=decy,y=S_abs_gsw))
  
#how else can you check data
hydrostation_bottle%>%
  filter(Sal1!=-999)%>%
  ggplot()+
  geom_point(aes(c=Sal1,y=S_abs_gsw))


#we need conservative temperature
#we need absolute salinity in situ-temp (ITS-90)

###calculate conserv temp
HydroS=
hydrostation_bottle=
  hydrostation_bottle%>%
  mutate(Pres_gsw=gsw_p_from_z(Depth*-1,latN))%>%
  mutate(s_abs_gsw=gsw_SA_from_SP(Sal1,Pres_gsw,360-lonW,latN))%>%
  mutate(T_cons_gsw=gsw_CT_from_t(S_abs_gsw,Temp,Pres_gsw))%>%
  mutate(Sig_th_gsw=gsw_sigma0(s_abs_T_cons_gsw))

HydroS%>%
  filter(`Sig-th`!=-999)%>%
  ggplot()+
  geom_point(aes(x=`Sig-th`,y=Sig_th_gsw))

HydroS %>% 
  filter(Sig_th_gsw<0) %>% 
  View()


## Homework
#something like select or combination of filters
#how ro replace Hydry a and b in a same argument.

         




  
```

