---
title: "TEOS-HydroStationS_Class"
author: Juanita Carballeira
output: html_document
date: '2023-03-18'
---

## Requiered Libraries
```{r, message=FALSE}
library(tidyverse)
#install.packages ("seacarb")
#Gibbs Seawater
library(seacarb)
library(gsw)
library(ggplot2)
```


## Import the data
```{r message=FALSE, warning=FALSE}
hydrostation_bottle <- read_delim("hydrostation_bottle.txt", 
    delim = "\t", escape_double = FALSE, 
    col_names = FALSE, trim_ws = TRUE, skip = 31)
hydrostation_bottle_names <- read_csv("hydrostation_bottle.txt", 
    skip = 30)
#Fix names for the dataset
colnames(hydrostation_bottle) = colnames(hydrostation_bottle_names)
view(hydrostation_bottle)
```
## Variables Names and Units
- yyyymmdd = Year Month Day   
- decy   = Decimal Year     
- time   = Time (hhmm)      
- latN   = Latitude (Deg N) 
- lonW   = Longitude (Deg W)
- Depth  = Depth (m)                  
- Temp   = Temperature ITS-90 (C) 
- Pres   = CTD Pressure (dbar)   
- CTD_S  = CTD Salinity (PSS-78)      
- Sal1   = Salinity-1 (PSS-78)        
- Sig-th = Sigma-Theta (kg/m^3) - density of seawater at a given temperature (Density - 1000)   
- O2(1)  = Oxygen-1 (umol/kg)          
- OxFixT = Oxygen Fix Temp (C)        
- Anom1  = Oxy Anomaly-1 (umol/kg)    

## Quality flags
*  -999  = No data
*  0 = Less than detection limit

```{r}
hydrostation_bottle %>% 
  filter(`Sig-th` != -999 & Depth < 20) %>% 
  ggplot()+
  geom_point(aes(x = decy, y = `Sig-th`))+
  geom_line(aes(x =decy, y = `Sig-th` ))
#Clear Seasonal signal for sigma-theta
hydrostation_bottle %>% 
  filter(`Sig-th` != -999 & Depth < 20) %>% 
  ggplot()+
  geom_point(aes(x =Temp, y = `Sig-th` ))
#Density and Temperature are strongly correlated, but there appears to be two outliners that we will likely need to address at some point
#We only have density data from 1988-present, but temperature and salinity data from 1950s-present, this mean i can calculate seawater density from 1950s- present, by using TEOS-10. 
```

## TEOS-10 Toolbox in package seacarb

```{r, include=FALSE, echo=FALSE}
#?gsw
#?gsw_sigma0 # It said we need absolute salinity and conservative temperature (SA (g/kg),CT (°C))
#?gsw_SA_from_SP #Absolute Salinity from Practical Salinity
#Practical Salinity*
#Sea pressure (dbar)
#Longitude
#Latitude
hydrostation_bottle %>% 
  ggplot()+
  geom_point(aes(x = decy, y = Pres ))
#Not for all the time series. Missing from 1980 
hydrostation_bottle %>% 
  ggplot()+
  geom_point(aes(x = decy, y = Depth ))
#We have Depth for the entire time series
#?gsw_p_from_z()
hydrostation_bottle =  hydrostation_bottle %>% 
  mutate(Pres_gsw = gsw_p_from_z(Depth*-1, latN))
hydrostation_bottle %>% 
  ggplot()+
  geom_point(aes(x = Pres, y = Pres_gsw))
# We see strong 1:1 agreement between measured pressure and calculated pressure
#Checking lat, lon, and salinity data
hydrostation_bottle %>% 
  ggplot()+
  geom_point(aes(x = decy, y = CTD_S))
hydrostation_bottle =  hydrostation_bottle %>% 
  mutate(Pres_gsw = gsw_p_from_z(Depth*-1, latN)) %>% 
  mutate(s_abs_gsw = gsw_SA_from_SP(Sal1, Pres_gsw,360-lonW, latN))
#Checking the Absolute salinity
hydrostation_bottle %>% 
  ggplot()+
  geom_point(aes(x = decy, y = s_abs_gsw))
#Plotting the ratio for Absoute Salinity (created) vs Sal1 (Given in the dataset)
hydrostation_bottle %>% 
  filter(Sal1!= -999) %>% 
  ggplot()+
  geom_point(aes(x = Sal1, y = s_abs_gsw))
#Temperature to Conservative Temperature
#?gsw_CT_from_t
HydroS =  hydrostation_bottle %>% 
  filter(Sal1!= -999) %>% 
  mutate(Pres_gsw = gsw_p_from_z(Depth*-1, latN)) %>% 
  mutate(s_abs_gsw = gsw_SA_from_SP(Sal1, Pres_gsw,360-lonW, latN)) %>% 
  mutate(T_Cons_gsw = gsw_CT_from_t(s_abs_gsw,Temp,Pres_gsw))
HydroS %>% 
  filter(Temp != -999) %>% 
  ggplot()+
  geom_point(aes(x = Temp, y = T_Cons_gsw))
HydroS =  hydrostation_bottle %>% 
  filter(Sal1 != -999 & `Sig-th` != -999) %>% 
  mutate(Pres_gsw = gsw_p_from_z(Depth*-1, latN)) %>% 
  mutate(s_abs_gsw = gsw_SA_from_SP(Sal1, Pres_gsw,360-lonW, latN)) %>% 
  mutate(T_Cons_gsw = gsw_CT_from_t(s_abs_gsw,Temp,Pres_gsw)) %>% 
  mutate(sig_theta_gsw = gsw_sigma0(s_abs_gsw,T_Cons_gsw))
HydroS %>% 
  filter(`Sig-th` != -999) %>% 
  ggplot()+
  geom_point(aes(x = `Sig-th`, y = sig_theta_gsw))
HydroS %>% 
  filter(sig_theta_gsw < 0) %>% 
  view()
```


```{r, include=FALSE, echo=FALSE}
HydroS_Corrected_a = HydroS %>% 
  filter(sig_theta_gsw<0) %>% 
  mutate(s_abs_gsw = gsw_SA_from_SP(CTD_S, Pres_gsw,360-lonW, latN)) %>% 
  mutate(T_Cons_gsw = gsw_CT_from_t(s_abs_gsw,Temp,Pres_gsw)) %>% 
  mutate(sig_theta_gsw = gsw_sigma0(s_abs_gsw,T_Cons_gsw)) 
  
HydroS_Corrected_b = HydroS %>% 
  filter(sig_theta_gsw>0) 
#The corrected one. And last one

HydroS_corrected = rbind(HydroS_Corrected_a,HydroS_Corrected_b) 
HydroS_corrected %>% 
  ggplot()+
  geom_point(aes(x = sig_theta_gsw, y = Depth))+
  scale_y_reverse()+
  scale_x_continuous(position = "top")+
  theme_classic()+
  xlab(expression(paste(sigma[theta], "(kg m"^"-3",")")))+
  ylab("Depth (m)")
                     
                     
```

#Has surface sigma theta decreased over time? 

```{r}
HydroS_Shallow = HydroS_corrected %>% 
  filter(Depth<30)
?lm
#lm(y is a function(~) of x)
lm(sig_theta_gsw~decy,data = HydroS_Shallow)
#y = mx+b
#Coeficients: intercept = b, decy = m 
#Sig_theta_gws = -0.01563*decy + 56.28
#(kg/m^3) = (kg/m^3/x)*x + (kg/m^3)
sig_theta_model = lm(sig_theta_gsw~decy,data = HydroS_Shallow)
summary(sig_theta_model)
library(plotly)
Sig_decy_plot = HydroS_Shallow %>% 
  ggplot(aes(x = decy, y = sig_theta_gsw))+
  geom_point()+
  geom_line()+
  geom_smooth(method = "lm")+
  theme_classic()
ggplotly(Sig_decy_plot)
```




## Lab Assigment
1. Pick a question (Include hypothesis)
Is density influenced more by temperature or salinity in the surface layer? Do salinity and temperature co-vary?

Hypothesis: Temperature have a greater effect on density in the surface layer.

2. produce a plot and a statistical summary
```{r}
Surface_hydro=HydroS_corrected%>%
  filter(Depth>100) 


Temp_densityplot=
  Surface_hydro%>%
  ggplot()+
  geom_point(aes(x=Depth,y= sig_theta_gsw))+
  xlab('Depth (m)')+
  ylab('Sig-theta (kg/m^3)')+
  scale_y_reverse()+
  scale_x_continuous(position ="top")

Temp_densityplot


```




3. Describe your results, the summary, and answer the question

4. Compile into a completed lab report using R Markdown

Potential questions:
how do temp, sal, and sigma-theta co-vary
Is there a relationship between sigma-theta and oxygen?
Is there a relationship of any of the parameter with depth? with time? within depth over time?
Are there seasonal differences in any of the parameters?

